Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-07-07 manuscript/manuscript.tex
Modifications to introduction

Lithosphere density variation with depth has been studied for almost a century. 
\citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Welford2010}.

These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation in regional scales.
This raises the importance of building forward models that reassemble the gravity anomalies for such scales. The main issue that should be overcome is that the curvature of the Earth must be taken into account, thus the modelling must be done in spherical coordinates.
One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids. \todo{Insertar figura de tesseroides de leo}
Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume elliptic integrals \todo{citar} that are generally approximated by numerical computations.
The literature offers two approaches: one involves Taylor series expansion \todo{citar} while the other makes use of Gauss-Legendre Quadrature (GLQ) \todo{citar}. The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.

@@

\citet{Uieda2016} developed a forward gravity model based on tesseroids with constant densities using the GLQ approximation method.
The GLQ method becomes less accurate when the computation point is closer to the tesseroid \todo{citar} and the smaller GLQ degree. 
Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation. % that has been objectively obtained by \citet{Uieda2016} comparing the analytical and the numerically approximated gravity fields generated by a spherical shell.
The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm instead of recursive function calls that  speeds up the algorithm and gives more control over the recursion step.
This forward gravity model has been originally implemented in command line programs written in C programming language. Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \todo{citar a fatiando} for geophysical modelling and inversion.

@@

>>>>

Lithosphere density variation with depth has been studied for almost a century. 
\citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.

These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation in regional scales.
This raises the importance of building forward models that reassemble the gravity anomalies for such scales. The main issue that should be overcome is that the curvature of the Earth must be taken into account, thus the modelling must be done in spherical coordinates.
One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids. \todo{Insertar figura de tesseroides de leo}
Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume elliptic integrals \todo{citar} that are generally approximated by numerical computations.
The literature offers two approaches: one involves Taylor series expansion \cite{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \cite{Asgharzadeh2007, Uieda2016, Uieda2017}. The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.

@@

\citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
The GLQ method becomes less accurate when the computation point is closer to the tesseroid \cite{Ku1977} or for smaller GLQ degree. 
Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.  \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm that  speeds up the computation and gives more control over the recursion step.

This forward gravity model has been originally implemented in command line programs written in C programming language. Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Theory}
